Sex-Based Differences in Tuberculosis Immune Response: The Role of Leptin

### 2025-04-14 products/manuscript/Manuscript.qmd
ML - part 4 - simulation

@@

@@

Our findings align with previous studies showing that females tend to have a stronger immune response than males [@Klein2016]. This difference in immunity may influence susceptibility and the speed of infection clearance between sexes [@Takahashi2021]. This study focused on immune activation and identified a sex-based difference in CD4+ cell levels, consistent with findings from a previous study [@Bothamley2021]. Recent research also indicates that changes in body fat in males can significantly reduce CD8+ cell levels compared to females [@Dhanyalayam2024]. Our study, along with previous research, provides evidence of sex-specific differences in the initiation and dynamic response of the immune system to TB infection. Our study also identified leptin as a bridge between metabolism and the immune system. This suggests its potential as a therapeutic target for regulating immune responses based on specific metabolic conditions [@Naylor2016]. Animal-based preclinical studies have demonstrated its practical significance [@Fantuzzi2000], but further research is needed to assess its clinical applicability in human populations.

>>>>

Last, we simulated a data set with log-transformed leptin level, sex, age, and BMI. This simulated data set was then used to examine the performance of random forest model in predicting leptin level using the other variables as predictors. The sex variable was generated using a Bernoulli distribution with a probability of 0.6 of being females. Age was assumed to follow a normal distribution with a mean of 35 and standard deviation of 3. BMI was also assumed a normal distribution with a mean of 25 and standard deviation of 1.5. We generated log-transformed leptin levels assuming the relationship between it and the other predictors as: $log-leptin=8*Sex-0.1*Age^2+6*Age+exp(0.05*BMI)$. Besides the association structure as described above, we also added an error term following a standard normal distribution. The model evaluation process was the same as conducted on the real-world data set.

@@

# Simulation study using random forest

The performance of the random forest model on the simulated data set appears to be highly accurate and optimized (@fig-rfsim, @fig-rfpredsim). The parameter tuning results demonstrate that the lowest RMSE was achieved when the number of randomly sampled predictors was set to 3 and the minimum number of split points was 37. In contrast, the model exhibited higher RMSEs as the number of randomly sampled predictors decreased, suggesting that incorporating a larger subset of predictors improves predictive performance. Additionally, the scatterplot comparing predicted and observed values shows that most points lie very close to the diagonal line, indicating strong agreement between predictions and actual outcomes. These findings collectively suggest that the random forest model effectively captured the complex, non-linear relationships embedded in the simulated data, resulting in robust and reliable predictions.

```{r}
#| label: fig-rfsim
#| fig-cap: "Parameter tuning results of random forest model using simulated data. "
#| echo: FALSE
knitr::include_graphics(here("results","figures","figure6.png"))
```

```{r}
#| label: fig-rfpredsim
#| fig-cap: "Fitting results of random forest on simulated data: predicted values vs. observed values. "
#| echo: FALSE
knitr::include_graphics(here("results","figures","figure7.png"))
```

@@

Our findings align with previous studies showing that females tend to have a stronger immune response than males [@Klein2016]. This difference in immunity may influence susceptibility and the speed of infection clearance between sexes [@Takahashi2021]. This study focused on immune activation and identified a sex-based difference in CD4+ cell levels, consistent with findings from a previous study [@Bothamley2021]. Recent research also indicates that changes in body fat in males can significantly reduce CD8+ cell levels compared to females [@Dhanyalayam2024]. Our study, along with previous research, provides evidence of sex-specific differences in the initiation and dynamic response of the immune system to TB infection. Our study also identified leptin as a bridge between metabolism and the immune system. This suggests its potential as a therapeutic target for regulating immune responses based on specific metabolic conditions [@Naylor2016]. Animal-based preclinical studies have demonstrated its practical significance [@Fantuzzi2000], but further research is needed to assess its clinical applicability in human populations. The extended two sections demonstrate the better performance of the random forest model in predicting leptin levels using body composition indices. Its strong performance is further demonstrated when non-linear relationships exist.
